---
title: "Natural Amenity Scale Ranking"
output:
  html_document:
    df_print: paged
---
This report shows some visualizations of the Natural Amenity Scale. 

```{r message=FALSE, echo=FALSE}
library(ggplot2)
setwd("~/Documents/__INFO-201/12code/nas")
source("./source/data_wrangling.R")
```

### The Data Set 
Recall that the data is organized in a table with a ranking assigned to each county in the U.S.A.

```{r echo=FALSE}
df <- create_natural_amenity_scale_df()
head(df)
```

### Question: What is the Distribution of Ranking Scores?

We can answer this question with a histogram.

```{r echo= FALSE, message=FALSE}
p <- qplot(data=df, x=Natural.Amenity.Rank)
print(p)
```

This is the call to qplot(). That's it! 

```{r eval=FALSE, echo= TRUE, message=FALSE}
qplot(data=df, x=Natural.Amenity.Rank)
```

We can add titles and labels to the graph, with this code.

```{r message=FALSE}
num <- nrow(df)
t <- paste0("Distribution of Natural Amenity Scale Rankings (N=",num,").")
p <- p + labs(title=t, 
     x = "Rank (1-low, 7-high)", 
     y = "Frequency")
print(p)
```

### Question: How are July and January Mean Temperatures Related?

We can answer this question with a scatterplot. Here's the code:

```{r message=FALSE}
p <- qplot(data=df, x=Mean.Temp.Jan, y=Mean.Temp.July, color=Natural.Amenity.Rank)
p <- p + labs(title="Natural Amenity Scale", 
     x = "Mean Temperature in January (Farienheit)", 
     y = "Mean Temperature in July (Farienheit)")
print(p)
```

### Question: Can we Discern Trends in this Data

This code can be used to create trend lines. Are there trends? 

```{r echo=FALSE, message=FALSE}
p <- qplot(data=df, x=Mean.Temp.Jan, y=Mean.Temp.July, color=factor(Natural.Amenity.Rank), alpha=0.25, geom=c("point", "smooth"))
p <- p + labs(title="Natural Amenity Scale", 
     x = "Mean Temperature in January (Farienheit)", 
     y = "Mean Temperature in July (Farienheit)")
print(p)
```

Here is the code:

```{r eval= FALSE, echo=FALSE, message=FALSE}
p <- qplot(data=df, x=Mean.Temp.Jan, y=Mean.Temp.July, color=factor(Natural.Amenity.Rank), alpha=0.25, geom=c("point", "smooth"))
p <- p + labs(title="Natural Amenity Scale", 
     x = "Mean Temperature in January (Farienheit)", 
     y = "Mean Temperature in July (Farienheit)")
print(p)
```

### Summary

qplot() is extremely effective for drawing standard charts. See the [reference guide](https://ggplot2.tidyverse.org/reference/qplot.html#arguments).
